Composable Generalized Consensus

### 2015-05-15 paper/paper.tex
Work on the description of the Simple-Leader algorithm.

\item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
from the acceptor $a$.

@@

\item The $Abort\left( s \right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
  at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message. The history $aborts'\left[ s \right]$ is then called an abort history. Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, the acceptor $a$ no longer accepts new proposals, therefore no history greater than $h$ can ever be learned anymore. Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.

\item The $Init\left( s \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$, and its effect is to pick a history $h$, set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.  The history $inits\left[ s \right]$ is then called an init history.
As described below, in $B_2$ composed with $B_1$, this history is constrained to be an abort history of $B_1$.

@@

The two rounds $B_1$ and $B_2$ are composed as follows:
\begin{itemize}
  \item We Require that a transition of the composition be a transition of either specification but not both, except for the $Abort\left( s \right)$ and $Init\left( s \right)$ transitions, which are allowed to take place at the same time. 
  \item We require that $aborts_1 = inits_2$ at all times.
\end{itemize}
The composition models two independent rounds being linked only by the switchers passing the abort histories from $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$, then they broadcast them as init histories to the acceptors in $B_2$. The acceptors then initialize their state to a received init history in the $WakeUp\left( a \right)$ transition.
This process can be repeated as often as needed, with new rounds $B_3$, $B_4$, etc., to try to make progress learning new histories.

@@

****************************************

@@

The specification of Simple-Leader algorithm uses the sets   
 $inits$, $aborts$, the arrays of variables $learned\left[ l \right]$,
$status\left[ a \right]$, and $accepted\left[ a \right]$, and two
additional arrays $accepted\left[ l \right]\left[ a \right]$ and
$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
an acceptor, and $s$ is a switcher.

Initially, the variables $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$ are empty sets. For every
acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
\right]$ is the special value $None$. For every learner $l$ and acceptor $a$,
$accepted\left[ l \right]\left[ a \right]$ is the special value
$None$. For every switcher $s$ and acceptor $a$, $stopped\left[ s
\right]\left[ a \right]$ is set to $false$.  There are initially no
messages in the network.

Thanks to the inter-round interface, Simple-Leader rounds and Broadcast-Decide rounds can be composed in any order.
To compose a Simple-Leader round $SL_1$ with an Broadcast-Decide round $B_2$, we compose their specifications as before, adding the constraint that the set $aborts_1$ be equal to the set $inits_2$ at all times.

>>>>

\item A $Learn\left( l, h \right)$ transition, for a learner $l$ and a history $h$, enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
from the acceptor $a$.
 Its effect is to set $learned\left[ l \right]$ to $h$.

@@

\item The $Abort\left( s, h\right)$ transition, for a switcher $s$, is enabled when $aborts\left[ s \right] = \quo{none}$ and the switcher has received
  at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the history $h$ received in the $\quo{stopped}$ message. The history $aborts'\left[ s \right]$ is then called an abort history. Because the switcher has received a $\aseq{\quo{stopped},a,h}$ message from the acceptor $a$, the acceptor $a$ no longer accepts new proposals, therefore no history greater than $h$ can ever be learned anymore. Therefore, an abort history can be seen as a safe starting point for a new round to resume execution.

\item The $Init\left( s,h \right)$ transition, for a switcher $s$, is enabled when $inits\left[ s \right]=\quo{none}$ and $h$ is a history, and its effect is to set $inits\left[ s \right]$ to $h$, and to broadcast the message $\aseq{\quo{init}, h}$ to the acceptors.  The history $inits\left[ s \right]$ is then called an init history.
As described below, in $B_2$ composed with $B_1$, this history is constrained to be an abort history of $B_1$.

@@

The composition $B_1\times B_2$ of the two rounds $B_1$ and $B_2$ is obtained as follows:
\begin{itemize}
  \item We Require that a transition of the composition be a transition of either specification but not both, except for the $Abort\left( s \right)$ and $Init\left( s \right)$ transitions, which are allowed to take place at the same time. 
  \item We require that $aborts_1 = inits_2$ at all times.
\end{itemize}
The composition models two independent rounds being linked only by the switchers passing the abort histories from $B_1$ to $B_2$, where they become init histories: switchers gather abort histories in $B_1$, then they broadcast them as init histories to the acceptors in $B_2$. The acceptors then initialize their state to a received init history in the $WakeUp\left( a \right)$ transition.
This process can be repeated as often as needed, with new rounds $B_3$, $B_4$, etc., to try to make progress learning new histories.

@@

@@

The specification of Simple-Leader algorithm uses the same variables as the Broadcast-Decide algorithm and the same initial state. 
%that is the arrays of variables    
%$inits\left[ s \right]$, $aborts\left[ s \right]$, $learned\left[ l \right]$,
%$status\left[ a \right]$, and $accepted\left[ a \right]$, where $l$ is a learner, $a$ is
%an acceptor, and $s$ is a switcher.

%Initially, the variables $inits$, $aborts$, and, for every
%learner $l$, $learned\left[ l \right]$ are empty sets. For every
%acceptor $a$, $status\left[ a \right]$ is ``idle'' and $hist\left[ a
%\right]$ is the special value $None$.

The next-state relation is the disjunction of the following types of
transitions. 
\begin{compactitem}
\item A $Propose\left( p,c \right)$ transition, for a proposer $p$
  and a command $c$, not guarded, sends the message $\aseq{\quo{prop},c}$ to
  the leader.
\item An $Init\left( s, h \right)$ transition, for a switcher $s$
  and a history $h$, is enabled when $inits\left[ s \right] = \quo{none}$, and its effect is to set $inits\left[ s \right]$ to $h$ and
  to sends the message $\aseq{\quo{init},h}$ to the leader.
\item A $WakeUp\left( leader \right)$ transition, enabled when
  $status\left[ leader \right]$ is ``idle'' and 
  a message $\aseq{\quo{init},h}$ was sent to the leader. Its effect is to
  receive the message, set $accepted\left[ leader \right]$ to $h$,
  to set $status\left[ leader \right]$ to ``ready'', and to
  broadcast the message $\aseq{\quo{leader-accept},h}$ to all
  other acceptors.
\item An $Accept\left( leader \right)$ transition, enabled when
  $status\left[ leader \right]$ is ``ready'' and a $\aseq{\quo{prop},c}$ message has been sent to the leader. The effect of the
  action is to receive the message, to update $accepted\left[ leader
  \right]$ to $accepted\left[ leader \right]\star c$, and to
  broadcast the message $\aseq{\quo{accepted},accepted'\left[ leader
  \right]}$ to the learners, and to broadcast the message
  $\aseq{\quo{leader-accept},accepted'\left[ leader \right]}$
  to all the other acceptors.
\item An $Accept\left( a \right)$ transition, for an acceptor $a$
  which is not the leader, is enabled when $status\left[ a
  \right]$ is ``ready'' and the leader has sent a
  $\aseq{\quo{leader-accept},h}$ message to $a$, where $h > accepted\left[ a
  \right]$. The effect of the action is to receive the message,
  to update $accepted\left[ a \right]$ to $h$, and to broadcast the
  message $\aseq{\quo{accepted},hist\left[ a \right]\star c}$ to the
  learners.
\item A $Learn\left( l, h \right)$ transition, for a learner $l$
  and a history $h$, is enabled when $h \geq learned\left[ l \right]$ and $h$ is a lower
  bound of the set $\aset{MaxReceivedFrom\left( a \right) : a \in Acceptors}$, where $MaxReceivedFrom\left( a \right)$ is the maximal accepted history received
  from the acceptor $a$.  Its effect is to set $learned\left[ l \right]$ to $h$.
\item A $Stop\left( s \right)$ transition, for a switcher $s$,
  enabled when the switcher $s$ has waited for too long without receiving any messages from the acceptors, broadcasts a $\aseq{\quo{stop}, s}$ message to all
  the acceptors.  
\item A $Stop\left( a \right)$ transition, for an acceptor $a$,
  enabled when $a$ can receive a $\aseq{\quo{stop},s}$ message
  from a switcher $s$. Its effect is to receive the message, set
  $status\left[ a \right]$ to ``stopped'', and to send the
  message $\aseq{\quo{stopped},a,hist\left[ a \right]}$ to the
  switcher $s$.
\item An $Abort\left( s, h \right)$ transition, for a switcher $s$ and a history $h$, is enabled when $aborts\left[ s \right] = \quo{none}$ and
  the switcher has received at least one $\aseq{\quo{stopped}, a, h}$ message. Its effect is to set the variable $aborts\left[ s \right]$ to the
  history $h$ received in the $\quo{stopped}$ message.
\end{compactitem}

Thanks to the inter-round interface, which is the same for Simple-Leader rounds and Broadcast-Decide rounds, Simple-Leader rounds and Broadcast-Decide rounds can be repeatedly composed in a sequence.
